---
output: github_document
---

# conpyro

<!-- badges: start -->

![lifecycle](https://img.shields.io/badge/lifecycle-beta-orange) ![license](https://img.shields.io/badge/license-GPL--3-blue)

<!-- badges: end -->

**Conifer Pyrometrics (ConPyro) fire behaviour modelling tools for R**

## Overview

`conpyro` is an R package implementing the **Conifer Pyrometrics (ConPyro)** fire behaviour modelling system described in:

> Perrakis, D. D. B., Cruz, M. G., Alexander, M. E., Hanes, C. C., Thompson, D. K., Taylor, S. W., & Stocks, B. J. (2023). Improved logistic models of crown fire probability in Canadian conifer forests. International Journal of Wildland Fire, 32(10), 1455–1473. <https://doi.org/10.1071/WF23074>

The package provides tools for estimating:

-   Fine fuel moisture content using a stand-adjusted model
-   Crown fire occurrence probability
-   Passive and active crown fire wind speed thresholds
-   Surface and crown fire rate of spread

The package is designed for both:

-   **Single-scenario exploratory modelling**, and
-   **Large multi-scenario workflows**, including modeling driven by spatial datasets, allowing users to evaluate fire behaviour across many fuel and fire weather configurations.

This GitHub release is a **beta version intended for testing and feedback prior to CRAN submission.**

**Note:** Both this software and the underlying ConPyro models are still under active development. Model outputs are estimates intended for informational and exploratory purposes only and should not be interpreted as definitive predictions of fire behaviour.

## Installation & Dependencies

Install the **beta version** from GitHub:

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("nhebda/conpyro")
```

Some optional helper functions rely on the package `cffdrs`. If you plan to use those tools, install it separately:

```{r eval=FALSE}
install.packages("cffdrs")
```

Optional plotting of results requires `ggplot2`:

```{r eval=FALSE}
install.packages("ggplot2")
```

## Basic Example

The main modelling function is `conpyro()`, which calculates crown fire occurrence probability, crowning thresholds, and rate of spread across a vector of 10-m open wind speeds.

```{r eval=FALSE}
conpyro(
  data = default_input,
  WS10 = 0:40,
  FFMC = 91,
  DMC = 85
)
```

This calculates fire behaviour metrics for each scenario defined in `data` across wind speeds from **0–40 km/h**.

## Input Methods

Inputs can be supplied in two ways:

### 1. Data frame input (recommended for many scenarios)

Each row in `data` represents a scenario – a single fuel and fire weather configuration. Thousands of scenarios can be quickly processed this way.

```{r eval=FALSE}
conpyro(
  data = my_scenarios,
  WS10 = 0:30
)
```

### 2. Direct arguments

Inputs may also be supplied by direct arguments, which can be be length `1` or the **common length of all scenarios**.

```{r eval=FALSE}
conpyro(
  WS10 = 0:40,
  FFMC = c(89, 91, 92),
  FSG = 6,
  SFC = 2,
  CBD = 0.1,
  DMC = c(23, 50, 80),
  season = c("spring", "summer", "summer"),
  density = "moderate",
  stand = "pine"
)
```

Length `1` arguments will be recycled across all scenarios.

## Plotting Results

`conpyro()` can optionally generate plots of model outputs.

```{r eval=FALSE}
conpyro(
  data = default_input,
  WS10 = 0:40,
  FFMC = 91,
  DMC = 85,
  plot = c("pCFO", "ROS")
)
```

Available plot types:

-   `"pCFO"` – Crown fire occurrence probability
-   `"CAC"` – Criterion for active crowning
-   `"ROS"` – Composite rate-of-spread plot with crowning thresholds

Plotting requires **ggplot2**.

## Helper Tools

The package includes a number of helper functions for calculating model inputs and intermediate quantities for single scenarios.

### Fine fuel moisture estimation

-   `t_mcF()` – FFMC-based fine fuel moisture content
-   `t_mcsa()` – Stand-adjusted fine fuel moisture

### Fire modelling

-   `t_pCFO()` – Crown fire occurrence probability
-   `t_FT()` – Fire type classification
-   `t_ROS()` – Equilibrium rate of spread

### Surface fuel consumption

-   `t_SFC_FBP()` – FBPS-based SFC
-   `t_SFC_deGroot()` – De Groot et al. (2009) SFC equation

### Foliar moisture

-   `t_FMC()` – Foliar moisture content estimate

### Experimental ladder fuel tools

-   `ladder_standing_dead()`
-   `ladder_midstory_saplings()`

These tools allow users to explore individual components of the ConPyro modeling system.

## Beta Status

This package is currently in **beta testing**.

During this phase we aim to:

-   Collect feedback from fire behaviour researchers and practitioners
-   Identify edge cases and bugs
-   Refine documentation and usability
-   Evaluate performance for large-scale modelling workflows

The package interface may evolve prior to a future CRAN release.

## License

GPL (\>= 3)

## Contact

Questions, bug reports, and feature requests are welcome.

Please open an issue on GitHub:

```         
https://github.com/nhebda/conpyro/issues
```

Or contact:

```         
Dr. Nicholas J. R. Hebda
Fire Informatics Officer
Canadian Forest Service
Natural Resources Canada
nicholas.hebda (at) nrcan-rncan.gc.ca
```
